---
title: "Lecture 5"
output: html_notebook
---

# Strength of seasonality and trend

```{r}
library(fpp3)
```

## Feature extraction and statistics

```{r}
tourism %>% 
  features(Trips, feat_stl)
```

```{r}
tourism %>% 
  features(Trips, feat_stl) %>% 
  ggplot(aes(x=trend_strength, y=seasonal_strength_year, col = Purpose)) +
  geom_point() +
  facet_wrap(vars(State))
```

Holidays more seasonal than other types (green dots high on y axis). Western Australia has the strongest trends. 

Find the most seasonal time series:

```{r}
most_seasonal <- tourism %>% 
  features(Trips, feat_stl) %>% 
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

```{r}
tourism %>% 
  right_join(most_seasonal, by = c("Region","State","Purpose")) %>% 
  ggplot(aes(Quarter, Trips)) + 
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

Same for most trended time series:
```{r}
most_trended <- tourism %>% 
  features(Trips, feat_stl) %>% 
  filter(trend_strength == max(trend_strength))

tourism %>% 
  right_join(most_trended, by = c("Region","State","Purpose")) %>% 
  ggplot(aes(Quarter, Trips)) + 
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

# Lab session 9

Use `GGally::ggpairs()` to look at the relationship between STL features. You might need to change `seasonal_peak_year` and `seasonal_trough_year` to factors.

```{r}
tourism %>% 
  features(Trips, feat_stl) %>% 
  select(-Region, -State, -Purpose) %>% 
  mutate(
    seasonal_peak_year = factor(seasonal_peak_year),
    seasonal_trough_year = factor(seasonal_trough_year)
  ) %>% 
  GGally::ggpairs()
```

This shows us where strange outlying values are. For isntance, very spikey series.
```{r}
tourism %>% 
  features(Trips, feat_stl) %>% 
  filter(spikiness > 2000)
```

Which is the peak quarter for travel in each state?
```{r}
tourism %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips)) %>% 
  features(Trips, feat_stl) %>% 
  select(State, seasonal_peak_year)
```

Features can be used for anomaloy detection across many time series.

## PCA on all feature sets

```{r}
# compute all features
tourism_features <- tourism %>% 
  features(Trips, feature_set(pkgs = "feasts"))

# pricipal components 
pcs <- tourism_features %>% 
  # remove non-numeric features
  select(-State, -Region, -Purpose) %>% 
  # scale to mean 0, var 1 and compute PCA
  prcomp(scale = TRUE) %>% 
  broom::augment(tourism_features)
```

```{r}
pcs %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2)) +
  geom_point() +
  theme(aspect.ratio = 1)
```

What's going on here?
```{r}
pcs %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = State)) +
  geom_point() +
  theme(aspect.ratio = 1)
```

Coloring by "State" doesn't help us. What about by "Purpose"?

```{r}
pcs %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, color = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
```

Filter for interesting outliers.
```{r}
outliers <- pcs %>% 
  filter(.fittedPC1 > 12 | .fittedPC1 > 10 & .fittedPC2 > 0)
outliers

outliers %>% 
  left_join(tourism, by = c("State","Region","Purpose")) %>% 
  mutate(Series = glue::glue("{State}","{Region}","{Purpose}",.sep="\n\n")) %>% 
  ggplot(aes(Quarter, Trips)) +
  geom_line() +
  facet_grid(Series ~ .) +
  labs(title = "outlying time series in the PCA space")
```


# Lab session 10

Use a features based approach to look for outlying series in PBS (Monthly Medicare Australia prescription data).

```{r}
autoplot(PBS, Cost) + theme(legend.position = "none")

pf <- PBS %>% features(Cost, feature_set(pkgs = "feasts"))
```

This won't work because two series have all zeros, and you can't compute models for data sets with all constants!
```{r}
# remove series with all zeros as features
PBS_no_zeros <- PBS %>% 
  group_by_key() %>% 
  filter(!all(Cost == 0)) %>% 
  ungroup()

# recompute PCs
pbs_features <- PBS_no_zeros %>% features(Cost, feature_set(pkgs = "feasts"))

# look for bad features caused by messy data that we didn't clean properly
summary(pbs_features)
```

```{r}
# compute PCs
pbs_prcomp <- pbs_features %>% 
  dplyr::select(-Concession, -Type, -ATC1, -ATC2) %>% 
  prcomp(scale = TRUE) %>% 
  broom::augment(pbs_features)

# plot first 2 components
pbs_prcomp %>% 
  ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point()

# pull out outliers
outliers <- pbs_prcomp %>% 
  filter(.fittedPC1 == max(.fittedPC1))
outliers
```

Visualize the outlying series.
```{r}
PBS %>% 
  semi_join(outliers, by = c("Concession", "Type","ATC1","ATC2")) %>% 
  autoplot(Cost) +
  facet_grid(vars(Concession, Type, ATC1, ATC2)) +
  theme(legend.position = "none")
```

This series has only two observations, so our `!any(x)` didn't catch and remove those series. However, using a features based appraoch, we were able to detect these anomalies without getting too deep into the data.



