---
title: "Lectures 3 and 4"
output: html_notebook
---

Consider the GDP information in `global_economy`. Plot the GSP per capita over time. Which country has the highest GDP per capita? How has this changed over time?

```{r}
library(fpp3)
```

```{r}
avg_gdp_ct <- global_economy %>% 
  as_tibble() %>% 
  group_by(Country, Year) %>% 
  mutate(gdp_per_capita = GDP/Population) %>% 
  ungroup()
  
avg_gdp_ct %>% 
  ggplot(aes(Year, gdp_per_capita, color = Country),
         alpha = 0.3) +
  geom_line() +
  guides(color = FALSE) +
  scale_color_viridis_d()
```

Top 5
```{r}
avg_gdp_ct %>% 
  arrange(desc(gdp_per_capita)) %>% 
  slice(1:5)
```


# Inflation adjustments

<see lecture notes PDF p 8-10>


# Mathematical adjustments

Variance stablization 

If the data show variation at different levels of the series, then a transformation can be useful. 

If $y_1 ... y_n$ are original, and $w_1 ... w_n$ are transformed, then by increasing strength:

* square root $w_t = \sqrt{y_t}$
* cube root $w_t = (y_t)^{1/3}$
* logarithm $w_t = log(y_t)$

Each of these transformations is close to a member of the family of Box-Cox transformations.

```{r}
food <- aus_retail %>% 
  filter(Industry == "Food retailing") %>% 
  summarise(Turnover = sum(Turnover))

autoplot(food)
```

```{r}
autoplot(food, sqrt(Turnover)) + labs(title="sqrt transformation")
autoplot(food, (Turnover)^(1/3)) + labs(title="cube root transformation")
autoplot(food, log(Turnover)) + labs(title="natural log transformation")
autoplot(food, (-1)/(Turnover)) + labs(title="inverse transformation")
```

In Box-cox transformations, $\lambda = 1$ corresponds to no substantive transformation, $\lambda = 0$ corresponds to a natural logarithm, and $\lambda = -1$ is an inverse plus 1 transformation.

We can find optimal values of $\lambda$ that stabalize variance via:
```{r}
food %>% features(Turnover, features = guerrero)
```

Always check results. A small $\lambda$ can give very large prediction intervals. 

```{r}
food %>% 
  autoplot(box_cox(Turnover, 0.0524)) +
  labs(title = "Box-Cox transformed turnover")
```

`logp1()` can be useful for data with zeros. 


# Lab session 7 

Find appropriate transformations to stablize variance of:

US GDP from `global_economy`

```{r}
us_gdp <- global_economy %>% 
  filter(Country == "United States")
us_gdp %>% features(GDP, features = guerrero)
```
Victoria bulls, bullocks steers 
```{r}
aus_livestock %>% 
  filter(
    State == "Victoria",
    Animal == "Bulls, bullocks and steers"
  ) %>% 
  autoplot(log(Count))
```

```{r}
vic_elec %>% autoplot(log(Demand))
```

```{r}
aus_production %>% autoplot(Gas)
aus_production %>% features(Gas, features = guerrero)
aus_production %>% autoplot(box_cox(Gas, 0.12))
```

In practice, going below 0 for $\lambda$ is very very uncommon.

***


The following is very hard to transform. Why?

The variation does not change with the level of the season. A transformation here isn't possible. Either need to only model the period after 1995, or accept a more uncertain model with large swings in seasonality.
```{r}
canadian_gas %>% autoplot()
```


# Time Series Decompositions

trend-cycle: aperiodic changes in level over time
seasonal: (almost) periodic changes in the level due to seasonal factors (e.g., quarter, month, year, etc...)

Additive decomposition:

$y_t = S_t + T_t + R_t$

$y_t$ is the data at time t  
$S, T, R$ are the seasonal, trend, and remainder components at time t  

## STL

STL: seasonal decomposition using loess

* additive model  
* seasonal component can change over time and specified by user  
* smoothness of trend cycle also controlled by user  
* optionally robust to outliers
* take logs for multiplicative decomposition
* box-cox for other trasformations

```{r}
ee <- as_tsibble(fpp2::elecequip)
dcmp <- ee %>% 
  model(STL(value ~ season(window = 7))) %>% 
  components()
dcmp
```

Top plot is an additive model of the three bottom plots.

```{r}
autoplot(dcmp) + 
  labs("STL decomposition", 
       subtitle = "value = trend + season_year + remainder")
```

```{r}
dcmp %>% gg_subseries(season_year)
```

Add a trendline
```{r}
ee %>% 
  autoplot(value) +
  autolayer(dcmp, trend, col="blue")
```

Recall holidays data
```{r}
holidays <- tourism %>% 
  filter(Purpose == "Holiday") %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))
holidays %>% autoplot()
```

Vertical rectangles are equal length scale bars to help visualize scales across axes of different limits.

```{r}
holidays %>% 
  model(stl = STL(Trips ~ season(window = "periodic"), robust = TRUE)) %>%
  components() %>% 
  autoplot()
```

Changing the window. Smaller windows are more smooth, larger values are more linear. Default values are sensible.
```{r}
holidays %>% 
  model(stl = STL(Trips ~ season(window = 5), robust = TRUE)) %>%
  components() %>% 
  autoplot()
```

```{r}
holidays %>% 
  model(stl = STL(Trips ~ season(window = 15) , robust = TRUE)) %>%
  components() %>% 
  autoplot()
```

Odd numbers are genearlly used because they contain the periods. 


# Lab session 8

Produce STL decomposition of Canadian gas volume:

```{r}
canadian_gas %>% autoplot()
```


```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 7) + trend(window = 11))) %>% 
  components() %>% 
  autoplot()
```

What happens when you change the values of the two window arguments?

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = "periodic") + trend(window = 11))) %>% 
  components() %>% 
  autoplot()
```

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 1) + trend(window = 11))) %>% 
  components() %>% 
  autoplot()
```

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 30) + trend(window = 11))) %>% 
  components() %>% 
  autoplot()
```

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 7) + trend(window = 100))) %>% 
  components() %>% 
  autoplot()
```

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 100) + trend(window = 100))) %>% 
  components() %>% 
  autoplot()
```

How does the seasonal shape change over time? 

```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 7) + trend(window = 11))) %>% 
  components() %>% 
  gg_season(season_year)
```

Can you produce a plausible seasonally adjusted series? [Hint: `season_adjust` is one of the variables returned by `STL`.]

Estimate seasonality with STL, remove seasonality, and return trend + remainder. This is what many economics statistics do when they "correct for seasonality".
```{r}
canadian_gas %>% 
  model(STL(Volume ~ season(window = 7) + trend(window = 11))) %>% 
  components() %>% 
  select(Month, season_adjust) %>% 
  autoplot()
```

# Multiple seasonality

```{r}
vic_elec %>% 
  model(STL(Demand)) %>% 
  components() %>% 
  autoplot()
```


# Seasonal adjustment

Useful byproduct of decomposition: easy way to calculate seasonally adjusted data.

Additive decompoosition of seasonally adjusted data:

$y_t - S_t = T_t + R_t$

Multiplicative decomposition of seasonally adjusted data:

$y_t / S_t = T_t * R_t$

Data in grey, seasonally adjusted data in blue.
```{r}
autoplot(ee, value, col = "gray") +
  autolayer(dcmp, season_adjust, col = "blue") +
  labs(title = "Electrical equipment manufacturing (Euro area)",
       x = "Year", y = "New orders index")
```



