---
title: "Lecture 10"
output: html_notebook
---

# Hierarchical and grouped time series

```{r}
library(fpp3)
```

Recall the tourism data set
```{r}
tourism
```

Region is nested with State. These variables are hierarchical. 

grouping factor: non-heirarchial (e.g., "Purpose")

## Creating aggregates

`aggregrate_key(State / Region, Trips = sum(Trips))` 

`aggregrate_key(Purpose * (State / Region), , Trips = sum(Trips))`

Useful when you want levels at different levels of aggregation. Simply adding up forecasts at lower levels does not always result in the same result as when you forecast at a higher level. Need to reconcile these forcasts with least squares. Implemented in `reconcile()`.




***

# Lab session 20

Prepare agregations of PBS data by Concession, Type, ATC1

```{r}
PBS_agg <- PBS %>% 
  aggregate_key(
    Concession * Type * ATC1,
    Cost = sum(Cost) / 1e6
  )

fit <- PBS_agg %>% 
  filter(Month <= yearmonth("2005 Jun")) %>% 
  model(
    ets = ETS(Cost),
    arima = ARIMA(Cost),
    snaive = SNAIVE(Cost)
  )

fc <- fit %>% 
  reconcile(
    ets_adj = min_trace(ets),
    arima_adj = min_trace(arima),
    snaive_adj = min_trace(snaive)
  ) %>% 
  forecast(h = "3 years") 

accuracy(fc, PBS_agg) %>% 
  group_by(.model) %>% 
  summarise(MASE = mean(MASE)) %>% 
  arrange(MASE)
```

