---
title: "Lecture 7"
output: html_notebook
---

```{r}
library(fpp3)
```

# Exponential smoothing

see pdf notes for details on additive and multiplicative models.

ETS(error, trend, season)

Each error, trend and season can be additive/multiplicative/none, or damped.

```{r}
aus_economy <- global_economy %>% 
  filter(Code == "AUS") %>% 
  mutate(Pop = Population/1e6)

fit <- aus_economy %>% 
  model(
    AAN = ETS(Pop)
  )

report(fit)
```

ETS(A,A,N) : A stands for "additive", M for "multiplicative", and N for "none". Damped is "Ad" or "Md" for additive and multiplicative respectively. 

$\sigma^2$ is the variance of the residuals

```{r}
components(fit)
components(fit) %>% autoplot()
```

```{r}
fit %>% 
  forecast(h = 20) %>% 
  autoplot(aus_economy)
```

Default chooses best (which is an AAN) in this case.
```{r}
aus_economy %>% 
  model(
    mod = ETS(Pop)
  ) %>% 
  report()
```

A damped example: it's leveling off. 
```{r}
fit <- aus_economy %>% 
  model(
    mod = ETS(Pop ~ trend("Ad"))
  ) 
report(fit)
autoplot(forecast(fit, h = 20), aus_economy)
```

An example with multiple series (and multiple models, best are selected automatically).
```{r}
fit <- global_economy %>% 
  mutate(pop = Population/1e6) %>% 
  model(ets = ETS(pop))

fit  

forecast(fit, h = 5)
```

# Lab session 14

Try foreasting Chinese GDP from global_economy using an ETS model.

```{r}
# calcualte gp per capita for china
china_gdp <- global_economy %>% 
  filter(Country == "China") 

autoplot(china_gdp, gdp_pc)

# fit ets model
china_gdp %>% 
  model(
    mod = ETS(GDP)
  ) %>% 
  forecast(h=10) %>% 
  autoplot(china_gdp)

# try some transformations
fit <- china_gdp %>% 
  model(
    ets = ETS(GDP),
    ets_damped = ETS(GDP ~trend("Ad")),
    # box cox tranformation via Guerro is actually negative, 
    # but never use soemthing that low in practice
    ets_bc = ETS(box_cox(GDP, 0.2)),
    ets_log = ETS(log(GDP))
  ) 

fit  

# ways to extract components
report(fit)
fit %>% select(ets) %>% report()
glance(fit)
tidy(fit)
coef(fit)

fit %>% 
  forecast(h="20 years") %>% 
  autoplot(china_gdp, level = NULL)
```

Example: Australian holiday tourism. Observe all the unique combinations of models generated. 
```{r}
holidays <- tourism %>% 
  filter(Purpose == "Holiday")
fit <- model(holidays, mod = ETS(Trips))
fit %>% pull(mod)
```

"Best" method chosen by AICc. See slide 39 of lecture 7. Paper for the method of "automatic forecasting" using AICc.

**ETS won't work well for data less than monthly level**. 


# Lab session 15

Find an ETS model for the Gas data from aus_prouduction

```{r}
autoplot(aus_production, Gas)

# needs a transformation
autoplot(aus_production, box_cox(Gas, 0.2))
autoplot(aus_production, log(Gas))

gas <- aus_production %>% 
  select(Gas) %>% 
  filter(Quarter <= yearquarter("2007 Q4")) 
autoplot(gas)

# fit models
fit <- gas %>% 
  model(
    auto = ETS(log(Gas)),
    damped = ETS(Gas ~ trend("Ad")),
    log = ETS(log(Gas))
  )

fit$auto    # AAA
fit$damped  # M Ad M
fit$log     # AAA

fc <- fit %>% 
  forecast(h = "4 years")

fc %>% 
  autoplot(aus_production, level = NULL)

# zoom in
fc %>% 
  autoplot(
    filter(
      aus_production, 
      Quarter > yearquarter("2000 Q4")),
    level = NULL)

fc %>% 
  accuracy(aus_production)

glance(fit)
```

```{r}
# note that the log transformed distributions are transformed normal distributions.
# it applies a bias correction factor in the backtransformation automatically. =)
fc %>% filter(.model == "log") %>% pull(.distribution)
```


# Non-Gaussian forecast distributions

If you don't want to believe normality. For instance, forecasts are too skewed (e.g. mean forecast not center of forecast distribution, or residuals not normally distributed).

For non-Gaussian resduals, you can bootstrap with `generate()`. Although this still assumed uncorrelated residuals. Can also do it within `forecast()` with `bootstrap = TRUE`. 500 samples is default number of samples, but you can changes this with `times = N`. See `?forecast.ETS()`.


# Lab session 

```{r}
vc <- aus_retail %>% 
  filter(State == "Victoria",
         Industry == "Cafes, restaurants and catering services") %>% 
  select(Month, Turnover)

vc %>% autoplot(Turnover)

# apply transformation to stabilize variance across time series
vc %>% autoplot(box_cox(Turnover, 0.2))

# fit model
fit <- vc %>% 
  model(ets = ETS(box_cox(Turnover, 0.2)))

fit$ets 
```

Bootstrapped forecast distributions
```{r}
sim <- fit %>% generate(h = "3 years", times = 5, bootstrap = TRUE)
sim

vc %>% 
  filter(Month >= 2008) %>% 
  ggplot(aes(x=Month)) + 
  geom_line(aes(y = Turnover)) +
  geom_line(data = sim, aes(y = .sim, color = as.factor(.rep))) +
  guides(col = FALSE) +
  coord_cartesian(xlim = yearquarter(c(2010, 2023)),
                  ylim = c(300, 800))
```



ETS handles trend, seasonality, but fails to handle cycles. For example:
```{r}
pelt %>% autoplot(Lynx)

# here's a lousy forecast
pelt %>% 
  model(ets = ETS(Lynx)) %>% 
  forecast(h="30 years") %>% 
  autoplot(pelt)

# observe the model: error is Additive, trend is none, season is none. 
# ETS model doesn't know what to do with cyclic data without seasonality!
pelt %>% 
  model(ets = ETS(Lynx)) %>% 
  pull(ets)
```

